---
title: "Lecture 6 R Functions"
author: "Simon Midtvedt"
date: "Jan 25th, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### File reading (again!)

Here we try to use **read.table()** and friends to input som example data into R.

```{r}
data1 <- read.csv("test1.txt")
data2 <- read.table("test2.txt", sep = "$", header = T)
data3 <- read.table("test3.txt")
```

## Hands-on Worksheet

### Section 1: Improving analysis code by writing functions
### **A.** Can you improve this analysis code?

```{r}
df <- data.frame(a=1:10, b=seq(200,400,length=10),c=11:20,d=NA)
df$a <- (df$a - min(df$a)) / (max(df$a) - min(df$a))
df$b <- (df$b - min(df$a)) / (max(df$b) - min(df$b))
df$c <- (df$c - min(df$c)) / (max(df$c) - min(df$c))
df$d <- (df$d - min(df$d)) / (max(df$a) - min(df$d))
```

Optimalization 1: 
```{r}
df <- data.frame(a=1:10, b=seq(200,400,length=10),c=11:20,d=NA)
normalize <- function(x){
  x <- (x - min(x)) / (max(x) - min(x))}
df$a <- normalize(df$a)
df$b <- normalize(df$b)
df$c <- normalize(df$c)
```

Optimalization 2:
```{r}
df <- data.frame(a=1:10, b=seq(200,400,length=10),c=11:20,d=NA)
normalize <- function(x){
  xmin <- min(x)
  x <- (x - xmin) / (max(x) - xmin)}
df$a <- normalize(df$a)
df$b <- normalize(df$b)
df$c <- normalize(df$c)
```

Optimalization 3:
```{r}
df <- data.frame(a=1:10, b=seq(200,400,length=10),c=11:20,d=NA)
normalize <- function(x){
  rng <- range(x)
  x <- (x - rng[1]) / (rng[2] - rng[1])}
df$a <- normalize(df$a)
df$b <- normalize(df$b)
df$c <- normalize(df$c)
```

Optimalization 4:
```{r}
df <- data.frame(a=1:10, b=seq(200,400,length=10),c=11:20,d=NA)
rescale <- function(x, na.rm = T, plot = F){
  if (na.rm) {
    rng <- range(x, na.rm = na.rm)
  } else {
    rng <- range(x)
  }
  print("Hello")
  
  answer <- (x - rng[1]) / (rng[2] - rng[1])
  
  print("is it me you are looking for")
  
  if(plot) {
    plot(answer, typ = "b", lwd = 4)
  }
  print("I can see it in...")
  
return(answer)
  }
```

Calling the finalized function:
```{r}
df$a <- rescale(df$a, plot = T)
df$b <- rescale(df$b, plot = T)
df$c <- rescale(df$c, plot = T)
df$d <- rescale(df$d, na.rm = F)
```

### **B.** Can you improve this analysis code?
```{r}
library(bio3d)

s1 <- read.pdb("4AKE")  # kinase with drug
s2 <- read.pdb("1AKE")  # kinase no drug
s3 <- read.pdb("1E4Y")  # kinase with drug

s1.chainA <- trim.pdb(s1, chain="A", elety="CA")
s2.chainA <- trim.pdb(s2, chain="A", elety="CA")
s3.chainA <- trim.pdb(s1, chain="A", elety="CA")

s1.b <- s1.chainA$atom$b
s2.b <- s2.chainA$atom$b
s3.b <- s3.chainA$atom$b

plotb3(s1.b, sse=s1.chainA, typ="l", ylab="Bfactor")
plotb3(s2.b, sse=s2.chainA, typ="l", ylab="Bfactor")
plotb3(s3.b, sse=s3.chainA, typ="l", ylab="Bfactor")
```

Optimalization 1:
```{r}
library(bio3d)

s1 <- read.pdb("4AKE")  # kinase with drug
s2 <- read.pdb("1AKE")  # kinase no drug
s3 <- read.pdb("1E4Y")  # kinase with drug

s1.chainA <- trim.pdb(s1, chain="A", elety="CA")
s2.chainA <- trim.pdb(s2, chain="A", elety="CA")
s3.chainA <- trim.pdb(s3, chain="A", elety="CA")

s1.b <- s1.chainA$atom$b
s2.b <- s2.chainA$atom$b
s3.b <- s3.chainA$atom$b

plotb3(s1.b, sse=s1.chainA, typ="l", ylab="Bfactor")
points(s2.b, sse=s2.chainA, col="red",typ="l", ylab="Bfactor")
plotb3(s2.b, sse=s2.chainA, typ="l", ylab="Bfactor")
plotb3(s3.b, sse=s3.chainA, typ="l", ylab="Bfactor")
```

Function:
```{r}
library(bio3d) # accessing the bio3d package

# Creating the function with inputs file (pdb structure file), chain (optional), elety (optional)
plotprot <- function(file, chain = "A", elety = "CA") { # chain and elety can be specified
  s <- read.pdb(file) # reads the protein databank file
  s.chain <- trim.pdb(s, chain = chain, elety = elety) # filters out the given chains and atoms in the pdb file using the trim function
  s.b <- s.chain$atom$b # selects the [b] vector from the [atom] dataframe within the selected chain. Bfactor represents isotropic displacement.
  
  plotb3(s.b, sse=s.chain, typ="l", ylab="Bfactor", main=file) # plot with secondary structure objects (sse)
}
```

Running the function with different proteins:
```{r}
plotprot("4AKE")
plotprot("1AKE")  
plotprot("1E4Y") 
```


