---
title: "Lecture 7 Functions and Packages"
author: "Simon Midtvedt"
date: "1/30/2019"
output: html_document
---

### Funcions revisited

Load (i.e. **source**) our rescale() funcion from last day.

```{r}
source("http://tinyurl.com/rescale-R")
```

Test this function

```{r}
rescale(1:5)
```

```{r}
rescale(c(1:5, "string"))
```

We want to make this function more 

```{r}
rescale2(c(1:5, "string"))
```

Now we want to make a function that finds NAs on the same spot in two vectors. We will start with forgetting about the function and focus on the task with an example.

```{r}
x <- c(1, 2, NA, 3, NA)
y <- c(NA, 3, NA, 3, 4)
```

Use Google to find a built-in function that may help

```{r}
is.na(x)
is.na(y)

is.na(x) & is.na(y)
sum(is.na(x) & is.na(y))
which(is.na(x) & is.na(y))
```

Now we take the working snippet and make a first function

```{r}
both_na <- function(x, y) {
  # Check for NA elements in both input vectors
  sum(is.na(x) & is.na(y))
}
```

```{r}
both_na(x, y)
```

```{r}
x <-  c(NA, NA, NA)
y1 <- c( 1, NA, NA)
y2 <- c( 1, NA, NA, NA)
y3 <- c(1, NA, NA, NA, NA)
```

```{r}
# What will this return?
both_na(x, y3)
```

```{r}
both_na2(x, y)
```

```{r}
both_na3(x, y)
```


